---
title: "read_me_game_data"
author: "Lucie Schulz"
date: '2024-03-19'
output: html_document
---

Title: FinSim Raw Game Data

Description:

This document explores the cleaned FinSim Game dataset. The data was collected during an RCT in Uganda in the summer of 2023 via a No-SQL Firestore database. The dataset contains a total of 117,950 decisions that were recorded from the game from a total of 133 unique participants, who played the game during the study. It tracks user activity within the FinSim Game, capturing details about each user's progress, decisions, and outcomes.

Source: The raw data is available via GitHub (ENTER URL)

Data Description:

RAW DATA COLLECTED IN REAL TIME DURING GAME PLAY
uid: Unique identifier of the participant
user_created_on: Date and time the user account was created.
session_id: Unique identifier for each game session.
session_started_on: Date and time the session began.
session_completed_on: Date and time the session ended (if applicable).
session_status: Status of the session (e.g., completed, abandoned).
level_id: Unique identifier for the game level played.
level_started_on: Date and time the level began within the session.
level_completed_on: Date and time the level was completed (if applicable).
level_status: Status of the level within the session (e.g. won, lost, incomplete).
level_number: Level number within the game (Level 1 - 6).
played_periods: Total number of periods played within a level (periods indicating investment rounds).
period_number: Period number within the level.
cash_at_start: User's starting cash balance for the period played.
cash_at_end: User's ending cash balance for the period played.
decision: User's decision made during a specific period (buyCash, loan, dontBuy).
income_per_period: Income earned per period from the presented asset.
life_expectancy: Asset's life expectancy within the game context.
price: Price of an asset within the game.
risk_level: Asset's risk level within the period (survival).
asset_type: Type of asset the user interacted with within the game (cow, chicken, goat).

CALCULATED FIELDS:
country: which country the study was conducted based on first two letters of uid (here UG = Uganda)
treatment_group: there were two treatment arms where the game was used (UGGM = game only and UGRT = game and aspiration treatment combined)
level_duration_minutes: how long a level was played (only available if completed or lost)
level_date: the date the level was played
decision_buycash: dummy variable created from decision variable in raw data
decision_dontbuy: dummy variable created from decision variable in raw data
decision_loan: dummy variable created from decision variable in raw data
expected_income_from_asset: Calculated by income_per_period * life expectancy
actual_price_asset: if the animal was bought via a loan = price * 1.2 (interest rate was set to 20% across the game)
roi_from_asset: expected income - actual price
should_buy_cash: TRUE if the ROI > 0
can_afford_in_cash: if start cash >= actual price of asset
can_afford_via_loan: if start case >= actual price of asset / 2 (loans repayed over 2 periods within the game)
should_borrow: TRUE if not enough cash and ROI of asset is positive even when loaning
type_one_error: User bought an asset with a negative ROI
type_two_error: User did not buy an asset with a positive ROI
correct_decision: Type 1 and Type 2 error both FALSE
session: in which session of the user the level was played (total of five sessions in study)

USER CHARACTERISTICS FROM SURVEY:
branch: which district the user is from
group: which micro-finance group the user is part of
gender: male / female
age: age group the user belongs to 
education: highest educational level reached by user
number_of_dependents: how many dependents the user has
income: Boolean, if income is generated
income_source1: what is the source of the income
digital_lit1: rating of digital literacy
fin_training: Boolean, if received financial training

Data Exploration:

```{r}
# libraries used
library(readxl)
library(kableExtra)

# Set the working directory - please update when using this script
setwd("/Users/lucieschulz/Desktop/Thesis/Data Analysis/cleaned_datasets/")

# Read the Excel file using the 'readxl' package
data <- read_csv("UG_cleaned_game_data.csv")

# Get the length (number of rows) of the DataFrame
length(data)

# unique participants
length(unique(data$uid))
```

Key Findings:

```{r}
# Summarize data
summary(data)
head(data)
```


